# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

```{r load_data}
activity_df <- read.csv(unz("activity.zip","activity.csv"))
activity_df$date <- as.Date(activity_df$date)
```

## What is mean total number of steps taken per day?
```{r mean_daily_steps}
library(dplyr)
library(ggplot2)

daily_totals <- activity_df %>%
                  group_by(date) %>%
                  summarise(total_steps = sum(steps))

ggplot(daily_totals, aes(x = total_steps)) +
      geom_histogram()

mean_steps <- mean(daily_totals$total_steps, na.rm = TRUE)
median_steps <- median(daily_totals$total_steps, na.rm = TRUE)
```
The **mean** number of steps per day is: `r mean_steps`

The **median** number of steps per day is: `r median_steps`

## What is the average daily activity pattern?
```{r ts_interval}
interval_mean <- activity_df %>%
                  group_by(interval) %>%
                  summarise(mean_steps = mean(steps, na.rm = TRUE))
ggplot(interval_mean, aes(x = interval, y = mean_steps)) +
      geom_line()

max_interval <- interval_mean$interval[which.max(interval_mean$mean_steps)]
max_steps <- max(interval_mean$mean_steps)
```
On average the `r max_interval` interval contains the **max** steps
of `r max_steps` steps.

## Imputing missing values
```{r count_nas}
summary(activity_df)
num_nas <- sum(is.na(activity_df$steps))
```
There are a total of `r num_nas` NA's in the dataset.

Use the average number of steps from each time interval to fill in NA's
```{r impute_nas}
na_inicies <- which(is.na(activity_df$steps))
na_steps_df <- activity_df[na_inicies,]
impute_vals <- left_join(x = na_steps_df, y = interval_mean)
steps_no_na <- activity_df$steps
steps_no_na[na_inicies] <- impute_vals$mean_steps
activity_df_clean <- activity_df
activity_df_clean$steps <- steps_no_na

summary(activity_df_clean)
```

```{r mean_daily_steps_clean}
daily_totals_clean <- activity_df_clean %>%
                          group_by(date) %>%
                          summarise(total_steps = sum(steps))

ggplot(daily_totals_clean, aes(x = total_steps)) +
      geom_histogram()

mean_steps_clean <- mean(daily_totals_clean$total_steps)
median_steps_clean <- median(daily_totals_clean$total_steps)
```
The **new mean** number of steps per day is: `r mean_steps_clean`

The **new median** number of steps per day is: `r median_steps_clean`

## Are there differences in activity patterns between weekdays and weekends?

```{r weekday vs weekend}
activity_df_clean$weekday <- weekdays(activity_df_clean$date)
activity_df_clean$day_category <- ifelse(activity_df_clean$weekday %in% c("Saturday","Sunday"),"Weekend", "Weekday")

weekday_interval <- activity_df_clean %>%
                        group_by(interval, day_category) %>%
                        summarise(mean_steps = mean(steps))

ggplot(weekday_interval, aes(x = interval, y = mean_steps)) +
      geom_line() +
      facet_grid(day_category~.)

```





